---
title: "Exploratory_Factor_Analysis"
output: html_document
date: "03/15/2023"
---

## Importing the complete college dataset
  
```{r}
library(psych)

clg <- read.csv("/Users//ajayvishnu/Desktop/RUTGERS/Spring 2023/Multivariate Analysis/Datasets/College_Data.csv", row.names=1)

attach(clg)
```

* Apps - Number of applications received
* Accept - Number of applications accepted
* Enroll - Number of new students enrolled
* Top10perc - Pct. new students from top 10% of H.S. class
* Top25perc - Pct. new students from top 25% of H.S. class
* F.Undergrad - Number of full-time undergraduates
* P.Undergrad - Number of part-time undergraduates
* Outstate - Out-of-state tuition
* Room.Board - Room and board costs
* Books - Estimated book costs
* Personal - Estimated personal spending
* PhD - Pct. of faculty with Ph.D.’s
* Terminal - Pct. of faculty with terminal degree
* S.F.Ratio - Student/faculty ratio
* perc.alumni - Pct. alumni who donate
* Expend - Instructional expenditure per student
* Grad_Rate - Graduation rate

## Using the varimax method

#### RC values, h2 & u2 values

```{r}
fit.pc <- principal(clg[-1], nfactors=4, rotate="varimax")
fit.pc
```

* h2 - Explained variance
* u2 - Unexplained variance
* h2 nd u2 add up to 1
* The Proportion explained shows us the PCA values

#### Eigen Values

```{r}
fit.pc$values
round(fit.pc$values, 3)
```

* The above are the eigen values.

#### Clubbing columns to RCs

```{r}
fit.pc$loadings
```

* We need to look for values over 0.5
* RC1 shows relation with Outstate, Expend, Top10perc, Room.Board, Grad.Rate, Top25perc, perc.alumni and negative with S.F.Ratio
* RC2 shows relation with Accept, Enroll, Apps, F.Undergrad, P.Undergrad
* RC4 shows relation with PhD and Terminal
* RC3 shows relation with Books and Personal

#### Loadings with more digits

```{r}
for (i in c(1,3,2,4)) { print(fit.pc$loadings[[1,i]])}
```

#### Communalities

```{r}
fit.pc$communality
```

#### Rotated factor scores, Notice the columns ordering: RC1, RC2, RC4 and RC3

```{r}
fit.pc$scores
```

#### Factor Recommendation

```{r}
fa.parallel(clg[-1])
```

* Plot shows elbow at 3

#### Correlations within Factors

```{r}
fa.plot(fit.pc)
```

#### Visualizing the relationship

```{r}
fa.diagram(fit.pc)
```

* RC1 - RC1 consisted of multiple aspects and is hard to interpret
* RC2 - RC2 summed up the applications, enrollements and the students aspects of the dataset.
* RC3 - RC3 summed up the expenses part in the dataset
* RC4 - RC4 summed up the faculty in the dataset

#### Factor recommendations for a simple structure

```{r}
vss(clg[-1])
```

* Till 3 there is change, from 3 no change
* This is how we decide number of factors

## Using oblimin rotate method

#### RC values, h2 & u2 values

```{r}
fit.pc1 <- principal(clg[-1], nfactors=4, rotate="oblimin")
fit.pc1
```

#### Eigen Values

```{r}
fit.pc1$values
round(fit.pc1$values, 3)
```

#### Clubbing columns to RCs

```{r}
fit.pc1$loadings
```

* We need to look for values over 0.5
* TC1 shows relation with Top10perc, Top25perc, Expend, PhD, Terminal, Outstate, Grade.Rate, Room.Board
* TC2 shows relation with F.Undergrad, Enroll, Accept, Apps, P.Undergrad, S.F.Ratio, and negative with perc.alumni
* TC3 shows relation with Books and Terminal

#### Loadings with more digits

```{r}
for (i in c(1,3,2,4)) { print(fit.pc1$loadings[[1,i]])}
```

#### Communalities

```{r}
fit.pc1$communality
```

#### Rotated factor scores, Notice the columns ordering: RC1, RC2, RC4 and RC3

```{r}
fit.pc1$scores
```

#### Correlations within Factors

```{r}
fa.plot(fit.pc1)
```

#### Visualizing the relationship

```{r}
fa.diagram(fit.pc1)
```

