---
title: "MVA_HW_1"
author: "aa2569@scarletmail.rutgers.edu"
date: "2/5/2023"
output: html_document
---

## Loading the Sparrows file 

```{r}
library(readr)
library(MVA)
library(HSAUR2)
Bumpus_sparrows <- read_csv("/Users/ajayvishnu/Desktop/Bumpus_sparrows.csv")
sparr <- Bumpus_sparrows
sparr
```

* We can see that the first column shows the Survivorship of the sparrows.
* NS - Not Survived, S - Survived.
* All the other 5 columns are numeric values.
* Later let us divide the data into two subsets (Survived & Not survived and perform the analysis on them separately)

## Separating into two different matrices

```{r}
ns_sparr <- subset(sparr, Survivorship == "NS")
s_sparr <- subset(sparr, Survivorship == "S")
ns_sparr
s_sparr
```

* The survived and not survived data has been seperated now.
* However, to perform the analysis, we need only the numeric data to be available.

## Droping the first column

```{r}
ns_sp <- ns_sparr[,-1]
s_sp <- s_sparr[,-1]
ns_sp
s_sp
```

* Further, we can calculate the column means, correlation, and covaraiance for the two datasets spearately and identify patterns.

## Column Means

```{r}
colMeans(ns_sp)
colMeans(s_sp)
```

* Except for L_humerous, rest all column means are larger for not survived sparrows.
* Further, we can deep dive and see the covariance and correlation matrices.

## Covariance Matrix

```{r}
cov(ns_sp)
cov(s_sp)
```

* The covariance of two random variables is a measure of their linear dependence.
* Larger values of the covariance imply a greater degree of linear dependence between two variables.
* For both not survived and survived, we can observe high covaraince between Total_Length and the Alar_extent.
* For both not survived and survived, we can observe least covaraince between L_humerous and the L_beak_head.
* However, the covariance values are higher for the not survived sparrows compared to survived sparrows.

## Correlation Matrix

```{r}
cor(ns_sp)
cor(s_sp)
```

* Compared to covariance, correlation gives us better understanding between the two variables.
* For not survived sparrows, we can see that the correlation is highest for the L_humerous and the L_beak_head and least for the L_keel_sternum and the L_beak_head.
* For survived sparrows, we can see that the correlation is highest for the L_humerous and the Alar_extent and least for the L_keel_sternum and the Alar_extent.
* We can see that these are different to that compared to our covaraince results.

## Distance Matrix

```{r}
dist_ns = dist(scale(ns_sp[, c("Total_length","Alar_extent","L_beak_head","L_humerous","L_keel_sternum")],center = FALSE))
dist_s = dist(scale(s_sp[, c("Total_length","Alar_extent","L_beak_head","L_humerous","L_keel_sternum")],center = FALSE))
dist_ns
dist_s
```

* The distance matrix gives us an idea of how like or unlike each pair of variables are in a dataframe.

## Not Survived Plots

```{r}
ns_sp <- as.data.frame(ns_sp)
cm_ns_sp <- colMeans(ns_sp)
cov_ns_sp <- cov(ns_sp)
d_ns <- apply(ns_sp, MARGIN = 1, function(ns_sp)t(ns_sp - cm_ns_sp) %*% solve(cov_ns_sp) %*% (ns_sp - cm_ns_sp))
qqnorm(ns_sp[,"Total_length"], main = "Total_length"); qqline(ns_sp[,"Total_length"])
qqnorm(ns_sp[,"Alar_extent"], main = "Alar_extent"); qqline(ns_sp[,"Alar_extent"])
qqnorm(ns_sp[,"L_beak_head"], main = "L_beak_head"); qqline(ns_sp[,"L_beak_head"])
qqnorm(ns_sp[,"L_humerous"], main = "L_humerous"); qqline(ns_sp[,"L_humerous"])
qqnorm(ns_sp[,"L_keel_sternum"], main = "L_keel_sternum"); qqline(ns_sp[,"L_keel_sternum"])
```

* We can see that the L_beak_head and L_keel_sternum are the only two which appear to be normally distributed with slight exclusions.
* The rest three are not normally distributed.

```{r}
plot(qc <- qchisq((1:nrow(ns_sp) - 1/2) / nrow(ns_sp), df = 5), sd <- sort(d_ns),xlab = expression(paste(chi[5]^2, " Quantile")),ylab = "Ordered distances")
oups <- which(rank(abs(qc - sd), ties = "random") > nrow(ns_sp) - 3)
text(qc[oups], sd[oups] - 1.5,oups)
abline(a = 0, b = 1)
```

* This distribution is also not normally distributed.
* There are also 3 outliers in the data which are observation 20, 24, and 28 in the ns_sp data frame.

## Survived Plots

```{r}
s_sp <- as.data.frame(s_sp)
cm_s_sp <- colMeans(s_sp)
cov_s_sp <- cov(s_sp)
d_s <- apply(s_sp, MARGIN = 1, function(s_sp)t(s_sp - cm_s_sp) %*% solve(cov_s_sp) %*% (s_sp - cm_s_sp))
qqnorm(s_sp[,"Total_length"], main = "Total_length"); qqline(s_sp[,"Total_length"])
qqnorm(s_sp[,"Alar_extent"], main = "Alar_extent"); qqline(s_sp[,"Alar_extent"])
qqnorm(s_sp[,"L_beak_head"], main = "L_beak_head"); qqline(s_sp[,"L_beak_head"])
qqnorm(s_sp[,"L_humerous"], main = "L_humerous"); qqline(s_sp[,"L_humerous"])
qqnorm(s_sp[,"L_keel_sternum"], main = "L_keel_sternum"); qqline(s_sp[,"L_keel_sternum"])
```

* For Survived sparrows, none of the 5 columns appear to be normally distributed.

```{r}
plot(qc1 <- qchisq((1:nrow(s_sp) - 1/2) / nrow(s_sp), df = 5), sd1 <- sort(d_s),xlab = expression(paste(chi[5]^2, " Quantile")),ylab = "Ordered distances")
oups <- which(rank(abs(qc1 - sd1), ties = "random") > nrow(s_sp) - 3)
text(qc[oups], sd[oups] - 1.5,oups)
abline(a = 0, b = 1)
```

* This distribution is also not normally distributed.
* There are also 3 outliers in the data which are observation 18, 19, and 20 in the s_sp data frame.
