---
title: "MVA_HW_2"
author: "aa2569@scarletmail.rutgers.edu"
date: "2/9/2023"
output: html_document
---

```{r}
library(readr)
library(SciViews)
library(scatterplot3d)
library(car)
library(lattice)
library(GGally)
library(ggplot2)
library(ggridges)
library(ggvis)
library(ggthemes)
library(cowplot)
library(gapminder)
library(gganimate)
library(dplyr)
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
sparrows <- read_csv("/Users/ajayvishnu/Desktop/RUTGERS/Spring 2023/Multivariate Analysis/Datasets/Bumpus_sparrows.csv")
str(sparrows)
boxplot(sparrows[,2:6])
stars(sparrows,labels = sparrows$Survivorship)
```

* The str function gives us a total summary of the data.
* The boxplot isn't giving much value as the data of different variables are at different levels.
* The stars function gave a much better result in this case. We can see that all the visuals for Survivors are all similar and the visuals for all non-survivors are all similar.

## Plots

```{r}
attach(sparrows)

sparr.birds <- data.frame(as.numeric(rownames(sparrows)),sparrows[,2:6])
labs.diagonal <- c("Bird","Total length","Alar extent","L. beak & head","L. humerous","L. keel & sternum")

plot(Total_length, Alar_extent,xlab="Total Length (mm)",ylab="Alar extent (nm)",pch=c(16,1))

pairs(sparrows[,2:6])
```

* Based on the pairs plot, we tried to check any correlations between the variables in the data.

```{r}
pairs(sparr.birds, diag.panel = panel.boxplot, labels=labs.diagonal,pch=c(1,16),font.labels=2)
```

* This gives us correlation plots along with boxplots for each variable.

## 3-D Plots

```{r}
s3d <- scatterplot3d(Alar_extent,Total_length,L_beak_head,pch=c(1,16)[as.numeric(Survivorship)],xlab="Alar extent", ylab="", angle=45,zlab="Length of beak and head", lty.hide=2,type="h",y.margin.add=0.1,font.axis=2,font.lab=2)
legend(s3d$xyz.convert(238, 160, 34.1),c("Non-survivor","Survivor"),pch=c(1,16),text.font=2)
```

## Scatterplot

```{r}
defaultW <- getOption('warn')
options(warn = -1)
scatterplotMatrix(~Total_length+Alar_extent+L_beak_head+L_humerous+L_keel_sternum | Survivorship, data=sparr.birds, var.labels=labs.diagonal,cex.labels=0.7, diagonal="boxplot",smooth=FALSE,reg.line=FALSE,pch=c(1,16),col=rep("black",2), legend.plot=FALSE)
options(warn = defaultW)
```

```{r}
super.sym <- trellis.par.get("superpose.symbol")
super.sym$superpose.symbol$pch <- c(1,16,rep(1,5))
super.sym$superpose.symbol$col <- rep("#000000",7)
trellis.par.set(super.sym)
splom(~sparr.birds, groups = Survivorship, data = sparr.birds, ps=0.5, varname.cex = .5,panel = panel.superpose,key = list(columns = 2,points = list(pch = super.sym$pch[1:2], col=super.sym$col[1:2]),text = list(c("Non-survivor", "Survivor"))))
```

```{r}
ggscatmat(sparrows, columns=2:6, color="Survivorship")
```

* All the above plots do not give us any proper inference that we can imply for sure.
* We further analyse using GGplots

## GG Plots

```{r}
ggplot(sparrows, aes(x=Survivorship,y=Total_length)) + geom_point(aes(col=Survivorship))
ggplot(sparrows, aes(x=Survivorship,y=Alar_extent)) + geom_point(aes(col=Survivorship))
ggplot(sparrows, aes(x=Survivorship,y=L_beak_head)) + geom_point(aes(col=Survivorship))
ggplot(sparrows, aes(x=Survivorship,y=L_keel_sternum)) + geom_point(aes(col=Survivorship))
ggplot(sparrows, aes(x=Survivorship,y=L_humerous)) + geom_point(aes(col=Survivorship))
```

```{r}
ggplot(sparrows, aes(x=Total_length,y=Survivorship)) + facet_wrap(Alar_extent) + geom_point()
ggplot(sparrows, aes(x=Total_length,y=Survivorship)) + facet_wrap(L_beak_head) + geom_point()
ggplot(sparrows, aes(x=Total_length,y=Survivorship)) + facet_wrap(L_humerous) + geom_point()
ggplot(sparrows, aes(x=Total_length,y=Survivorship)) + facet_wrap(L_keel_sternum) + geom_point()
```

```{r}
ggplot(sparrows, aes(Total_length, col = Survivorship)) + geom_bar(position="stack") 
ggplot(sparrows, aes(Alar_extent, col = Survivorship)) + geom_bar(position="stack") 
ggplot(sparrows, aes(L_beak_head, col = Survivorship)) + geom_bar(position="stack") 
ggplot(sparrows, aes(L_humerous, col = Survivorship)) + geom_bar(position="stack") 
ggplot(sparrows, aes(L_keel_sternum, col = Survivorship)) + geom_bar(position="stack") 
```

```{r}
ggplot(sparrows, aes(Total_length, col=Survivorship)) + facet_grid(.~Survivorship) + geom_bar(position="dodge")
ggplot(sparrows, aes(Alar_extent, col=Survivorship)) + facet_grid(.~Survivorship) + geom_bar(position="dodge")
ggplot(sparrows, aes(L_beak_head, col=Survivorship)) + facet_grid(.~Survivorship) + geom_bar(position="dodge")
ggplot(sparrows, aes(L_humerous, col=Survivorship)) + facet_grid(.~Survivorship) + geom_bar(position="dodge")
ggplot(sparrows, aes(L_keel_sternum, col=Survivorship)) + facet_grid(.~Survivorship) + geom_bar(position="dodge")
```

```{r}
ggplot(sparrows, aes(Total_length))+geom_histogram()
ggplot(sparrows, aes(Alar_extent))+geom_histogram()
ggplot(sparrows, aes(L_beak_head))+geom_histogram()
ggplot(sparrows, aes(L_humerous))+geom_histogram()
ggplot(sparrows, aes(L_keel_sternum))+geom_histogram()
```

```{r}
ggplot(sparrows, aes(Total_length))+geom_histogram(aes(fill = after_stat(count)))
ggplot(sparrows, aes(Alar_extent))+geom_histogram(aes(fill = after_stat(count)))
ggplot(sparrows, aes(L_beak_head))+geom_histogram(aes(fill = after_stat(count)))
ggplot(sparrows, aes(L_humerous))+geom_histogram(aes(fill = after_stat(count)))
ggplot(sparrows, aes(L_keel_sternum))+geom_histogram(aes(fill = after_stat(count)))
```

```{r}
ggplot(sparrows, aes(x=Survivorship, y=Total_length)) + geom_violin()
ggplot(sparrows, aes(x=Survivorship, y=Alar_extent)) + geom_violin()
ggplot(sparrows, aes(x=Survivorship, y=L_beak_head)) + geom_violin()
ggplot(sparrows, aes(x=Survivorship, y=L_humerous)) + geom_violin()
ggplot(sparrows, aes(x=Survivorship, y=L_keel_sternum)) + geom_violin()
```

```{r}
ggplot(sparrows, aes(x=Survivorship, y=Total_length)) + geom_boxplot()
ggplot(sparrows, aes(x=Survivorship, y=Alar_extent)) + geom_boxplot()
ggplot(sparrows, aes(x=Survivorship, y=L_beak_head)) + geom_boxplot()
ggplot(sparrows, aes(x=Survivorship, y=L_humerous)) + geom_boxplot()
ggplot(sparrows, aes(x=Survivorship, y=L_keel_sternum)) + geom_boxplot()
```

```{r}
ggplot(sparrows, aes(x=Total_length, fill=Survivorship, color=Survivorship)) + geom_density(alpha=0.3, aes(y=..scaled..)) 
ggplot(sparrows, aes(x=Alar_extent, fill=Survivorship, color=Survivorship)) + geom_density(alpha=0.3, aes(y=..scaled..)) 
ggplot(sparrows, aes(x=L_beak_head, fill=Survivorship, color=Survivorship)) + geom_density(alpha=0.3, aes(y=..scaled..)) 
ggplot(sparrows, aes(x=L_humerous, fill=Survivorship, color=Survivorship)) + geom_density(alpha=0.3, aes(y=..scaled..)) 
ggplot(sparrows, aes(x=L_keel_sternum, fill=Survivorship, color=Survivorship)) + geom_density(alpha=0.3, aes(y=..scaled..)) 
```

```{r}
ggplot(sparrows, aes(x=Total_length, y=Survivorship)) + geom_density_ridges() 
ggplot(sparrows, aes(x=Alar_extent, y=Survivorship)) + geom_density_ridges() 
ggplot(sparrows, aes(x=L_beak_head, y=Survivorship)) + geom_density_ridges() 
ggplot(sparrows, aes(x=L_humerous, y=Survivorship)) + geom_density_ridges() 
ggplot(sparrows, aes(x=L_keel_sternum, y=Survivorship)) + geom_density_ridges() 
```

```{r}
ggplot(sparrows, aes(x=Survivorship, y=Total_length)) + geom_hex()
ggplot(sparrows, aes(x=Survivorship, y=Alar_extent)) + geom_hex()
ggplot(sparrows, aes(x=Survivorship, y=L_beak_head)) + geom_hex()
ggplot(sparrows, aes(x=Survivorship, y=L_humerous)) + geom_hex()
ggplot(sparrows, aes(x=Survivorship, y=L_keel_sternum)) + geom_hex()
```

```{r}
lastplot <- ggplot(sparrows, aes(x=Alar_extent,y=L_keel_sternum)) + xlim(220,260) + geom_point(aes(color=Survivorship)) + 
  labs(x="Alar_Extent", y="L_keel_sternum", title="Sparrows Analysis") 

lastplot + theme_bw()
lastplot + theme_cowplot()
lastplot + theme_dark()
lastplot + theme_economist()
lastplot + theme_fivethirtyeight()
lastplot + theme_tufte()
lastplot + theme_wsj()
```

* Based on all the plots, we can infere that the birds with lengths of the variables in the extremes did not survive.
* For all the birds that survived, they had the lengths in the average range. 
* For those that did not survive, all the birds which had the lengths in the extremes have died for sure.
* However, we cannot confidently say a range in which the birds will die or survive.









